---
title: "Homework 8"
output: html_notebook
---

#### Rice 11.35
An experiment was done to measure the effects of ozone, a component of smog.
A group of 22 seventy-day-old rats were kept in an environment containing ozone
for 7 days, and their weight gains were recorded. Another group of 23 rats of a
similar age were kept in an ozone-free environment for a similar time, and their
weight gains were recorded. The data (in grams) are given below. Analyze the
data to determine the effect of ozone. Write a summary of your conclusions.
[This problem is from Doksum and Sievers (1976) who provide an interesting
analysis.]

```{r}
control = c(41.0,25.9,13.1,-16.9,15.4,22.4,29.4,26.0,38.4,21.9,27.3,17.4,27.4,17.7,21.4,26.6,24.9,18.3,28.5,21.8,19.2,26.0,22.7)
ozone = c(10.1,7.3,-9.9,17.9,6.6,39.9,-14.7,-9.0,6.1,14.3,6.8,-12.9,12.1,-15.9,44.1,20.4,15.5,28.2,14.0,15.7,54.6,-9.0)
```

Find the mean and the standard error.

```{r}
control_bar = mean(control)
ozone_bar = mean(ozone)
sprintf("mean of control: %s, mean of ozone : %s", control_bar, ozone_bar)
```

```{r}
control_var = var(control)
ozone_var = var(ozone)
sprintf("var of control: %s, var of ozone : %s", control_bar, ozone_var)
```

Find the pooled sample variance.

```{r}
n = length(control)
m = length(ozone)
pooled_var = ((n - 1) * control_var + (m - 1) * ozone_var) / (n + m - 2)
sprintf("pooled variance: %s", pooled_var)
```

Calculate the test statistic.

```{r}
test_statistic = (control_bar - ozone_bar) / sqrt(pooled_var * (1 / n + 1 / m))
sprintf("test statistic: %s", test_statistic)

p = 2 * (1 - pnorm(test_statistic))
sprintf("p-value: %s", p)
```

Reject the null hypothesis that there is no effect.


#### Rice 11.36
Lin, Sutton, and Qurashi (1979) compared microbiological and hydroxylamine
methods for the analysis of ampicillin dosages. In one series of experiments, pairs
of tablets were analyzed by the two methods. The data in the following table give
the percentages of claimed amount of ampicillin found by the two methods in
several pairs of tablets. What are X − Y and s X −Y ? If the pairing had been erro-
neously ignored and it had been assumed that the two samples were independent,
what would have been the estimate of the standard deviation of X − Y ? Ana-
lyze the data to determine if there is a systematic difference between the two
methods.

```{r}
micro = c(97.2,105.8,99.5,100.0,93.8,79.2,72.0,72.0,69.5,20.5,95.2,90.8,96.2,96.2,91.0)
hydro = c(97.2,97.8,96.2,101.8,88.0,74.0,75.0,67.5,65.8,21.2,94.8,95.8,98.0,99.0,100.2)
```

Calculate mean, sample variances and the test statistic, if the data is pair-wise.

```{r}
diff = micro - hydro
diff_bar = mean(diff)
diff_var = var(diff)
test_statistic = diff_bar / sqrt(diff_var / length(diff))

test_statistic
```

Calculate p-value.

```{r}
p = 2 * (1 - pnorm(test_statistic))
p
```

Do the same calculations if it is independent.

```{r}
n = length(micro)
pooled_var = (var(micro) + var(hydro)) / 2
test_statistic = (mean(micro) - mean(hydro)) / sqrt(2 * pooled_var / n)
p = 2 * (1 - pnorm(test_statistic))
p
```

You would not reject it in either case, so there is no significant difference.

#### Rice 11.37
Stanley and Walton (1961) ran a controlled clinical trial to investigate the effect
of the drug stelazine on chronic schizophrenics. The trials were conducted on
chronic schizophrenics in two closed wards. In each of the wards, the patients were
divided into two groups matched for age, length of time in the hospital, and score
on a behavior rating sheet. One member of each pair was given stelazine, and the
other a placebo. Only the hospital pharmacist knew which member of each pair
received the actual drug. The following table gives the behavioral rating scores
for the patients at the beginning of the trial and after 3 mo. High scores are good.

a. For each of the wards, test whether stelazine is associated with improvement
in the patients’ scores.

Data for patients with Stelazine and placebo before and after.
```{r}
sb = c(2.3,2.0,1.9,3.1,2.2,2.3,2.8,1.9,1.1)
sa = c(3.1,2.1,2.45,3.7,2.54,3.72,4.54,1.61,1.63)
pb = c(2.4,2.2,2.1,2.9,2.2,2.4,2.7,1.9,1.3)
pa = c(2.0,2.6,2.0,2.0,2.4,3.18,3.0,2.54,1.72)

s_diff = sa - sb
p_diff = pa - pb
```

```{r}
n = length(s_diff)
pooled_var = (var(s_diff) + var(p_diff)) / 2
test_statistic = (mean(s_diff) - mean(p_diff)) / sqrt(2 * pooled_var / n)
p = 2 * (1 - pt(test_statistic, df = 2 * n - 2))

p
```

Would not reject with 95% confidence interval. Let's verify using R's built-in t-test.

```{r}
t.test(s_diff, p_diff, alternative = 'less')
```

#### Rice 11.39
An experiment was done to test a method for reducing faults on telephone lines
(Welch 1987). Fourteen matched pairs of areas were used. The following table
shows the fault rates for the control areas and for the test areas:

```{r}
test = c(676,206,230,256,280,433,337,466,497,512,794,428,452,512)
control = c(88,570,605,617,653,2913,924,286,1098,982,2346,321,615,519)
```


a. Plot the differences versus the control rate and summarize what you see.

```{r}
plot(test)
points(control, pch = 5)
legend(1, 95, legend = c('test', 'control'), col = c('red', 'blue'))
```
Circle is test, diamond is control.

b. Calculate the mean difference, its standard deviation, and a confidence interval.

```{r}
diff = test - control
mean = mean(diff)
var = var(diff)
sprintf("mean: %s, sample var: %s", mean, var)
```

```{r}
alpha = 0.95
t = qt((1 + alpha) / 2, df = length(diff))
me = t * sqrt(var / length(diff))
lower = mean - me
upper = mean + me
sprintf("%s confidence interval for the change is (%s, %s)", alpha, lower, upper)
```

We reject that there is not a significant difference between the new method and the control group.